\vspace{-6.0\baselineskip}
\textbf{\begin{center}
		{\large \textbf{ACKNOWLEDGEMENTS}}
\end{center}}
\vspace{0.1\baselineskip}
\noindent\rule{\linewidth}{2pt}


It is a great pleasure for me to express my respect and deep sense of gratitude to my supervisor \textbf{Dr. A. R. Jayan}, Professor, Department of Electronics \& Communication Engineering, Government Engineering College, Thrissur, for his vision,  guidance, enthusiastic involvement and persistent encouragement during the planning and development of this research work.

I make use of this opportunity to thank my co-supervisor \textbf{Dr. Rajeev Rajan}, Associate Professor, Department of Electronics \& Communication Engineering, Government Engineering College, Bartonhill, Thiruvananthapuram for his expertise, wisdom, and involvement  without which this work could not have been completed. 

I would like to extend my heartfelt appreciation to the members of the Doctoral Committee for their valuable feedback and constructive suggestions, which have been  helpful in shaping the direction of my research. I am forever thankful to my supervisors, for their trust and support, which provided me with the freedom to explore and experiment without constraints. Their guidance was instrumental in staying focused to my research goals.

My sincere gratitude goes to the members of the \textbf{Swathanthra Malayalam Computing}, whose efforts on creating accessible language technology tools has motivated me to ensure that my research and its outcomes are available to the public under permissive licenses.

I am immensely grateful to my fellow researchers, \textbf{Amlu}, \textbf{Lekshmi}, \textbf{Noumida}, and \textbf{Mala}, for their invaluable support and willingness to go the extra mile to help in times of need. I fondly acknowledge the encouragement, timely help and cooperation, I have received from the staff, faculty members and fellow researchers at CET, during various stages of this research. 

I express my gratitude to my parents and my sister, whose enthusiastic interest in my work has been a source of motivation for me and I am honoured to have them as my family. I would like to express my heartfelt gratitude and love to \textbf{Santhosh}, whose presence in my life has been a spirited source of strength. I am beyond words to thank his dedication to our growth and well-being. I thank \textbf{Vani}, for joining us in the middle of this journey and filling our life with love, warmth, and joy. 


% Their consistent encouragement and inquiries about my progress have helped me stay focused and committed to my goals.


% have been a constant source of motivation for me, and I am thankful for their presence in my life. Their belief in me and my abilities has been instrumental in helping me reach this point, and I am honored to have them as my family."

% I thank them from the bottom of my heart for being pillars of strength throughout this journey.
 
 


% \par I am obliged to my parents \textcolor{blue}{Father name} and \textcolor{blue}{Mother name} for their moral support, love, encouragement and blessings to complete this task. I am especially thankful to my wife \textcolor{blue}{Wife name} for her patience, love and encouragement during this journey.

% \par I am also thankful to the anonymous reviewers of my research publications. Their comments and suggestions were very helpful in shaping my research work.

% I would also like to thank DRC members for their constructive suggestions, support and encouragement.
% \par I would also like to extend my special thanks to \textcolor{blue}{XYZ} and other staff members of computer center in ECE Department, for their timely help and cooperation extended throughout the course of investigation.
% \par Finally, I am indebted and grateful to the Almighty for helping me in this endeavor.




\vspace{1cm}
\begin{flushright}
	\textbf{\thescholar}
\end{flushright}

